Functions and Relations

### 2023-10-22 rec08/paper.tex
functions

\begin{lem}
  Composition is associative, i.e., $(f \circ g) \circ h = f \circ (g \circ h)$.
\end{lem}
\begin{proof}
  Exercise.
  Hint: Use functional extensionality.
\end{proof}

@@

\begin{lem}
  For any function $f : A \to B$, $\id_{B} \circ f = f$ and $f \circ \id_{A} = f$.
\end{lem}
\begin{proof}
  Exercise.
\end{proof}

@@

>>>>

\begin{lem}\label{lem:associativity}
  Composition is associative, i.e., $(f \circ g) \circ h = f \circ (g \circ h)$.
\end{lem}
\begin{proof}
  Exercise.
  Hint: Use functional extensionality.
\end{proof}

@@

\begin{lem}\label{lem:unitality}
  For any function $f : A \to B$, $\id_{B} \circ f = f$ and $f \circ \id_{A} = f$.
\end{lem}
\begin{proof}
  Exercise.
\end{proof}

\cref{lem:associativity,lem:unitality} mean that sets and functions between them assemble into a category.
Category theory is an interesting subject that we will sadly not discuss in this class.

@@

\begin{defn}
  A function $f : A \to B$ is \emph{invertible} if there is a function $g : B \to A$ such that
  \begin{enumerate}
  \item $f \circ g = \id_{B}$, and
  \item $g \circ f = \id_{A}$.
  \end{enumerate}
  $g$ is called the inverse of $f$.
  When $f$ is invertible, we write $f\inv$ for its inverse.
\end{defn}

\begin{thm}
  A function $f : A \to B$ is invertible if and only if $f$ is bijective.
\end{thm}
\begin{proof}
  The ``if'' direction: assume that $f$ is invertible.
  Then there is a function $f\inv : B \to A$ such that $f \circ f\inv = \id_{B}$ and $f\inv \circ f = \id_{A}$.

  (Injectivity): Let $a, a' \in A$ be given.
  Assume that $f(a) = f(a')$.
  Then $\id_{A}(a) = f\inv(f(a)) = f\inv(f(a')) = \id_{A}(a')$.
  Thus, $a = a'$.

  (Surjectivity): Let $b \in B$ be given.
  We need to show that there is some $a \in A$ so that $f(a) = b$.
  Choose $a := f\inv(b)$, then $f(f\inv(b)) = \id_{B}(b) = b$.

  The ``only if'' direction: assume that $f$ is bijective.
  We need to show that $f$ is invertible.
  To this end, we construct a function $f\inv : B \to A$.
  For each $a \in A$ so that $f(a) = b$, we take $f\inv(b) = a$.
  To show that $f\inv$ is a function, we must show that it is total and functional.
  Totality follows from surjectivity of $f$ and functionality follows from injectivity of $f$.
  The details are left to the reader as an exercise.
\end{proof}
